---
title: "Configurations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Configurations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Livestock movement and holding data can come in a diversity of shapes and formats. To ensure 
correct reading of data regardless of the format, movenet requires its users to provide
configurations, indicating how the package should read any incoming movement and holding data files.

This vignette describes how to manage movenet configurations.

```{r setup}
library(movenet)
```

## Configuration files

Configuration (config) files tell movenet how to read incoming data files. 

There are two types of config files, each with slightly different requirements: 

- movement config files, with information on how to read movement data files. Configuration groupings in these files are prefixed with `movedata_`.
- holding config files, with information on how to read holding data files. Configuration groupings in these files are prefixed with `holdingdata_`.

Movement and holding configurations can also be combined into a single config file, containing both `movedata_` and `holdingdata_` configuration groupings. These combined config files need to fulfill criteria for both movement and holding config files. 

While differing in the details, movement and holding config files share a common structure, built around
two groups of configurations: 

- file and data options that commonly vary between datasets (`fileopts`)
- column headers/indices for data fields that are either required or that you want to extract along for analyses 
(`cols`)

These groupings are visualised in the content of the example movement config file `ScotEID.yml`, and discussed in more detail in the subsections below.
```{R}
xfun::file_string(system.file("configurations", "ScotEID.yml", package="movenet"))
```

Various example config files and empty templates can be found in movenet's `configurations`
folder. 


### File and data options

Configurations in the file and data options grouping (`movedata_fileopts` or `holdingdata_fileopts`) literally tell movenet how to read a data file.

The following file and data options are required, and config files will not be validated or 
loaded if these configurations are absent:

- `separator`: separator (or delimiter) character used in the data file
- `encoding`: encoding used in the data file
- `decimal`: decimal mark used in the data file
- `date_format`: date format specification used in the data file (required for movement config files only)

Additionally, holding config files may contain several file and data options that are required if you wish to extract and process columns with geographical coordinates:

- `coord_EPSG_code`: numeric part of the EPSG code for the Coordinate Reference System used in the data file
- `country_code`: two-letter country code for the data in the data file

### Column headers/indices

Configurations in the column headers/indices grouping (`movedata_cols` or `holdingdata_cols`) indicate columns to be extracted from the data files. They can be provided as column headers (character) or as column indices (integer).

For movement config files, the following column headers/indices are required, and movement config files will not be validated or loaded if these configurations are absent:

- `from`: header/index of the column containing identifiers of the holdings of origin
- `to`: header/index of the column containing identifiers of the destination holdings
- `date`: header/index of the column containing movement dates
- `weight`: header/index of the column containing movement weights/quantities

For holding config files, there is only one required column header/index, and movement config files will not be validated or loaded if this configuration is absent:

- `id`: header/index of the column containing holding identifiers, matching to `from`/`to` in movement data

Additionally, holding config files may contain several optional column headers/indices relating to holding properties that are commonly relevant for analyses:

- `coord_x`: header/index of the column containing longitudinal geographical coordinates (x/longitude) of holdings
- `coord_y`: header/index of the column containing latitudinal geographical coordinates (y/latitude) of holdings 
- `herd_size`: header/index of the column containing herd sizes

Furthermore, in both movement and holding config files, additional column headers/indices can be included, to indicate optional columns you wish to extract along with the required columns. These optional configurations can be given any name.

## Loading a configuration file
The function `load_config()` first validates a config file, and then loads its configurations 
into the movenet environment.

Use `load_config(configfile)`, with `configfile` consisting of either the path to a 
config file, or the name (minus extension) of a pre-installed config file.

```{R}
# Load a pre-installed config file by using just its name:
load_config("ScotEID")
# Alternatively, provide the path to any configurations file:
load_config(system.file("configurations", "fakeScotEID_holding.yml", package="movenet"))
```

The movenet environment can hold one set of movement configurations and one set of holding 
configurations at the same time.

## Querying configurations
The function `get_config()` can be used to view the configurations currently loaded 
in the movenet environment.

To view specific configurations, use `get_config(...)`, with the value of each argument 
consisting of a character string combining the following elements: 

- `"movedata_"` or `"holdingdata_"`
- `"fileopts"` for file options, or `"cols"` for column headers
- a full-stop followed by the configuration name, e.g. `".separator"`.

Multiple configurations can be queried at once:
```{R}
# Query the values of specific configurations: 
get_config("movedata_fileopts.separator", "holdingdata_cols.id")
```

Alternatively, to get a full list of loaded configurations, use `get_config()` without arguments:
```{R}
# Query the values of all configurations: 
get_config()
```

## Saving configurations to a config file
The function `save_config()` saves the currently loaded configurations to a new config 
file.

It requires the argument `outfile`, being the file name (and path) to save 
the config file to. It also takes an argument `config_type` that indicates which 
configurations you would like to save: accepted values are `"movement"`, `"holding"`, 
or `c("movement", "holding")` (default).

```{R eval = FALSE}
# Save the currently loaded movement configurations to a new config file:
save_config("saved_movement_config.yml", config_type = "movement")
```

## Copying a config file template to your working directory
The function `new_config()` copies a config file template to your working directory. 

It takes an argument `config_type` that indicates which config file template you would 
like to copy: accepted values are `"movement"`, `"holding"`, or `c("movement", "holding")` 
(default).

```{R eval = FALSE}
# Copy the movement config file template to your working directory:  
new_config(config_type = "movement")
```

## Changing configurations
The function `change_config()` can be used to change individual configurations in the 
movenet environment. 

Configuration names and new values should be passed to `change_config(...)` in `name = value` 
format, with each configuration name consisting of a character string combining 
the following elements: 

- `"movedata_"` or `"holdingdata_"`
- `"fileopts"` for file options, or `"cols"` for column headers
- a full-stop followed by the configuration name, e.g. `".separator"`.

Multiple configurations can be changed at once:
```{R}
# Change the values of specific configurations: 
change_config("movedata_fileopts.separator" = ";", "holdingdata_cols.id" = "foo")
# Inspect the changed values in the movenet environment:
get_config("movedata_fileopts.separator", "holdingdata_cols.id")
```

## Setting a default configuration file

If you use movenet with the same configuration most or all of the time, you may 
want to have movenet automatically load a configuration file when the movenet package 
is loaded.

To do this, you should find the `movenet/R` directory (where the package's main 
code is located). Create a file in this directory called `zzz.R`. In this file, 
type the following code:

```{r eval=FALSE}
.onLoad <- function(libname, pkgname){
  load_config("ScotEID")
}
```

where `ScotEID` can be replaced with whatever configuration you wish
to load. Now, whenever the movenet package is loaded, this `.onLoad`
function will be called, loading the desired configuration. See
[this page](https://r-pkgs.org/Code.html#sec-code-onLoad-onAttach)
for more information on `.onLoad` and `zzz.R`.
