---
title: "Getting started with movenet"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with movenet}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal of movenet is to simplify the use of livestock movement data in
veterinary public health. It does this by providing workflows for
pseudonymisation, network analysis, and integration of movement and/or
holding data into transmission models. A first step in any of these
workflows is reading in the livestock movement data that you want to
process or analyse.

This vignette describes how to get started with movenet, by reading in
data files and reshaping them into a common format required by all three
movenet workflows.

```{r setup}
library(movenet)
```

## Reshaping livestock movement data to movenet format

To read in a movement data file, first load a configurations (config)
file, telling movenet how to read the data file, and then load the data
file itself.

### Loading a movement config file

Livestock movement data can come in a diversity of shapes and formats:
not only do different countries collect different types of trade-related
data, they may also follow different data formatting customs for, for
example, decimal marks and dates.

To ensure correct reading of the data regardless of the source, movenet
requires users to supply a config file. The config file contains values
for a range of properties that commonly vary between datasets, as well
as column headers for data fields that are either required (`from`, `to`,
`date`, `weight`) or that you want to extract along for analyses. For
details on config file requirements and how to manage movenet
configurations, see `vignette("configurations")`. Example config files
and an empty template can be found in the package's `configurations`
folder.

Here is the content of an example config file, `ScotEID.yml`:

```{R}
xfun::file_string(system.file("configurations", "ScotEID.yml", package="movenet"))
```

The `ScotEID.yml` configurations can be loaded into the movenet
environment with the function `load_config()`:

```{R}
load_config(system.file("configurations", "ScotEID.yml", package="movenet")) # Load ScotEID.yml

get_config() # Inspect the configurations in the movenet environment
```

### Loading a movement data file

After configurations have been loaded into the movenet environment, a
delimited movement data file can be read in and processed.

Here are the first 6 rows of an example movement data file, `fake_Scottish_movement_data.csv`:

```{R}
head(read.csv(system.file("extdata", "fake_Scottish_movement_data.csv", package="movenet"), 
              encoding = "UTF-8"))
```

The function `reformat_data()` uses the loaded configurations to read in
the movement data, and standardise them to movenet format, by:

-   extracting columns with minimally required data, corresponding to `from` (originating holding identifier), `to` (destination holding identifier), `date`, and `weight` (movement quantity). The resulting data tibble contains these columns in this specific order.

-   extracting along any optional columns, as indicated
    in the loaded configurations. In the resulting data tibble, these
    columns are located *after* the required data columns.

-   checking that the `date` column can be parsed as dates (using the date format 
    indicated in the configurations), and that the `weight` column is numeric.

-   converting dates to R Date format, to improve interoperability.


To read in a movement data file, use `reformat_data(data, type = "movement")`:
```{R}
movement_data <- 
  reformat_data(system.file("extdata", "fake_Scottish_movement_data.csv", package="movenet"),
                type = "movement")

head(movement_data) # Inspect the resulting movement_data tibble
```

The movement data are now in the right format (movenet-format movement data tibble) 
to be plugged into pseudonymisation, network analysis, and transmission modelling workflows.

## Reshaping holding data to movenet format

If you want to use additional holding data (e.g. coordinates or holding type) in 
your analyses, the same process as described above can be followed for holding 
config and data files:
```{R}
# Load a holding config file:
load_config(system.file("configurations", "fakeScotEID_holding.yml", package="movenet")) 
# Read in and reformat a holding data file:
holding_data <- 
  reformat_data(system.file("extdata", "fake_Scottish_holding_data.csv", package="movenet"),
                type = "holding")
```

For holding data, the only absolutely required data column is `id` (holding identifier, 
matching to `from` and `to` in the movement data). If you wish to include geographic 
coordinates, `coord_EPSG_code` and `country_code` are required file options, and `coord_x` 
and `coord_y` are required data columns.

In addition to extraction of required and requested optional columns, `reformat_data()` 
performs the following data checks and standardisations on holding data:

-   If the loaded configurations include headers/indices for columns with geographical 
    coordinates (`coord_x` and `coord_y`), these columns are checked to contain 
    numeric data, and are then converted to a single simple feature (sf) geometry 
    list-column using the ETRS89 coordinate reference system.
    
-   If the loaded configurations include a header/index for a column `herd_size`, this 
    column is checked to contain numeric data.
    
The holding data are now in the right format (movenet-format holding data tibble) 
to be plugged into pseudonymisation, network analysis, and transmission modelling 
workflows.
