---
title: "Network analysis of livestock movement data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Network analysis of livestock movement data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
movenet uses the [Statnet suite of packages](https://statnet.org/packages/) (e.g. networkDynamic and tsna) for (temporal) social network analysis. We are currently working on  the functions in this section - e.g. creation of a pdf report with network measures of particular relevance to disease transmission.

```{r setup}
library(movenet)

# Load example movenet-format movement and holding tibbles into the global environment:
movement_data <- head(fake_Scottish_movement_data, 100)
holding_data <- fake_Scottish_holding_data
```

## Generating a temporal network representation

Currently it is possible to generate a temporal network (networkDynamic) representation
of movement data and optional holding data, as follows:

```{R warning = FALSE}
# Create a networkDynamic from our data
network <- movedata2networkDynamic(movement_data, holding_data,
                                   incl_nonactive_holdings = FALSE)
network
```

## Calculate summary statistics for selected temporal network measures

```{R}
# Find the maximum reachability in the network, and which nodes have this property.
parallel_summarise_temporal_node_properties(list(network), n_threads = 1,
                                            node_property = "forward reachability",
                                            statistics = list(max = max),
                                            identify_nodes = TRUE)
```


## Create a temporal social network analysis report

```{R, eval = FALSE}
# Create a network analysis report, with analyses of monthly network snapshots
create_temporal_network_analysis_report(
  network,
  output_file = paste0(tempdir(),"\\network_report.yml"),
  incl_reachability_analysis = FALSE,
  n_threads = 1,
  whole_months = TRUE,
  time_unit = "month")
  
```

