---
title: "Pseudonymising livestock movement and holding data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Pseudonymising livestock movement and holding data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

To help address concerns around commercial sensitivity, movenet includes a range of 
functions to make livestock movement data and/or holding data non-identifiable. 
This vignette shows you how to use these.

```{r setup}
library(movenet)

# Load a combined movement and holding config file:
load_config(system.file("configurations", "fakeScotEID_combined_config.yml", package="movenet"))

# Load example movenet-format movement and holding tibbles into the global environment:
data(example_movement_data, package = "movenet")
data(example_holding_data, package = "movenet")
```

## Pseudonymising holding identifiers

The function `anonymise()` pseudonymises holding identifiers in movenet-format 
movement or holding data tibbles, by replacing these identifiers with a number
and an optional prefix (e.g. "FARM").

It returns a pseudonymised data tibble and the applied pseudonymisation key. This 
key can optionally be saved to recover the original identifiers at a later date, or for 
application to an overlapping dataset.
```{R}
# Pseudonymise movement_data by changing identifiers to FARM1-N:
pseudonymised <- anonymise(example_movement_data, prefix = "FARM")
pseudonymised_movement_data <- pseudonymised$data
pseudonymisation_key <- pseudonymised$key

head(pseudonymised_movement_data) # Inspect pseudonymised movement data
head(pseudonymisation_key) # Inspect pseudonymisation key
```

`anonymise()` also takes an optional `key` argument, with which you can apply an existing
pseudonymisation key to the data tibble:

```{R}
# Use the same key from above to substitute holding identifiers in holding_data:
pseudonymised_holding <- anonymise(example_holding_data, key = pseudonymisation_key) 
pseudonymised_holding_data <- pseudonymised_holding$data
# Update saved key, in case additional identifiers were added from the holding datafile:
pseudonymisation_key <- pseudonymised_holding$key 

head(pseudonymised_holding_data) # Inspect pseudonymised holding data
```

This allows multiple datasets to be pseudonymised in a consistent way, so that
it is possible to subsequently merge the datasets by pseudonymised identifier.

## Modifying dates, weights, and optional numeric data columns

movenet also has functions to modify movement dates or weights by applying a small amount 
of noise (jittering) or by rounding:

- `jitter_dates(data, range)` adds random noise of up to `range` days to movement dates.

- `jitter_weights(data, range, column)` adds random noise of up to `range` to a numeric 
`column` in the movement data, by default the "weight" column.

- `round_dates(data, unit, week_start, sum_weight, ...)` rounds movement dates down to 
the first day of the specified time `unit`. For rounding down to weeks, set the starting 
day of the week with `week_start`. By default, weights are aggregated for all 
movements between the same holdings over the indicated time unit (`sum_weight = TRUE`); 
to keep movements separate, set `sum_weight = FALSE`. Alternative or additional summary 
functions can be applied through `...`, using tidy evaluation rules.

- `round_weights(data, unit, column)` rounds data in a numeric `column`, by default 
the "weight" column, to multiples of `unit`.

```{R}
# Add jitter of up to ±5 days to movement dates:
movedata_datesj5 <- jitter_dates(example_movement_data, range = 5) 
head(example_movement_data) # Inspect original
head(movedata_datesj5) # Inspect jittered dates

# Add jitter of up to ±10 to movement weights:
movedata_weightsj10 <- jitter_weights(example_movement_data, range = 10)
head(movedata_weightsj10) # Inspect jittered weights

# Round movement dates down to the first day of the month, but do not aggregate:
movedata_months <- round_dates(example_movement_data, unit = "month", sum_weight = FALSE) 
head(movedata_months) # Inspect rounded dates

# Round movement dates down to the first day of the month, aggregate weights, and list reference numbers:
movedata_months_aggr <- round_dates(example_movement_data, unit = "month", sum_weight = TRUE,
                                    movement_reference = list(movement_reference))
# Inspect aggregated record for holdings which have 2 movements in the same month:
movedata_months_aggr[which(sapply(movedata_months_aggr$movement_reference, length) == 2),] 

# Round movement reference numbers to the nearest multiple of 10:
movedata_ref10 <- round_weights(example_movement_data, unit = 10, column = "movement_reference")
head(movedata_ref10) # Inspect rounded movement reference numbers

```


## Modifying holding coordinates

A function to resample holding coordinates in a density-dependent manner is under 
development within the [hexscape package](https://github.com/ku-awdc/hexscape/).
