---
title: "Temporal network analysis"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
```

This page follows analyses and recreates some plots from [Schulz J et
al. (2017) PLOS ONE 12(6):
e0179915](https://doi.org/10.1371/journal.pone.0179915).

# Basic network summary

## Network size {.tabset}

**Overall network size:** `r n_nodes` holdings active over
the observed time period

<div class = "row">
<div class = "col-md-5">
```{r, network-size-table, echo= FALSE}
qflextable(size_summary_table)
```
</div>
<div class = "col-md-7">
### Holdings active (n)
```{r, network-size-plot-n, echo= FALSE}
plot(size_plot_number)
```

### Holdings active (prop)
```{r, network-size-plot-prop, echo= FALSE}
plot(size_plot_prop)
```
</div>
</div>

## Movements and edges {.tabset}

**Overall number of movements:** `r n_moves` movements over the observed
time period  
**Overall number of edges:** `r edge_count` different links
between holdings over the observed time period  
**Overall edge density:** `r round(edge_density, 3)` of all possible links between holdings are
observed over the observed time period

<div class = "row">
<div class = "col-md-5">
```{r, moves-summary-table, echo= FALSE}
qflextable(moves_summary_table)
```
</div>
<div class = "col-md-7">
### Movements
```{r, moves-plot-n, echo= FALSE}
plot(moves_plot_number)
```

### Movements (prop.)
```{r, moves-plot-prop, echo= FALSE}
plot(moves_plot_prop)
```

### Edges
```{r, edges-plot-n, echo= FALSE}
plot(edges_plot_number)
```

### Edges (prop.)
```{r, edges-plot-prop, echo= FALSE}
plot(edges_plot_prop)
```

### Density
```{r, edge-density-plot, echo= FALSE}
plot(edges_plot_density)
```
</div>
</div>

## Movement weights (batch sizes)

<div class = "row">
<div class = "col-md-5">
```{r, batch-sizes-table, echo= FALSE}
qflextable(weights_summary)
```
</div>
<div class = "col-md-7">
```{r, batch-sizes-plot, echo= FALSE}
plot(weights_plot)
```
</div>
</div>

# Static network analysis (`r time_unit`ly snapshots)

## Component analysis

<div class = "row">
<div class = "col-md-7">
```{r, GSCC-component-plot, echo= FALSE}
plot(GSCC_size_plot)
```
</div>
<div class = "col-md-5">
<br>
<br>
The largest giant strongly connected component (GSCC) is the largest
number of nodes (holdings) that can all reach each other by some
directed path of movements. This means that any node in that set could
indirectly infect any other, or that a single source outbreak within the
component could theoretically infect anything else within that
component.  
</div>
</div>

<div class = "row">
<div class = "col-md-7">
```{r, GIC-component-plot, echo= FALSE}
plot(GIC_size_plot)
```
</div>
<div class = "col-md-5">
<br>
<br>
The giant in-component (GIC) consists of all nodes (holdings) that can reach the 
GSCC, but are not part of it, because there is no directed path from the GSCC to 
these nodes. This means that any holding in this set could infect holdings in 
the GSCC, but not be infected by them. 
</div>
</div>

<div class = "row">
<div class = "col-md-7">
```{r, GOC-component-plot, echo= FALSE}
plot(GIC_size_plot)
```
</div>
<div class = "col-md-5">
<br>
<br>
Conversely, the giant out-component (GOC) consists of all nodes (holdings) that 
can be reached from the GSCC, but are not part of it, because there is no 
directed path from these nodes to the GSCC. This means that any holding in this 
set could be infected by holdings in the GSCC, but not infect them.
</div>
</div>


## Loyalty

<div class = "row">
<div class = "col-md-7">
```{r, loyalty-plot, echo= FALSE}
plot(common_links_heatmap)
```
</div>
<div class = "col-md-5">
<br>
<br>
This is a measure of consistency in the network between different points
in time. Darker colours indicate a larger proportion of trades that are
in common between two months. The black diagonal indicates that each
month's movements are identical to themselves.
</div>
</div>


```{r, reachability-analysis, echo= FALSE, fig.show="hold", out.width="50%", eval = incl_reachability_analysis, results='asis'}

cat("# Temporal network analysis\n\n")
cat("## Reachability (in- & out-going contact chains)\n\n")

plot(fwd_reachability_plot)
plot(bkwd_reachability_plot)

```

`r if(incl_reachability_analysis){"The forward reachable set (or out-going contact chain) of a single node
is the set of nodes that can be reached from that origin node, or
alternatively the set of nodes that could theoretically be infected in
an outbreak starting at that node. That is, the set of nodes that are
reachable from some origin node are the reachable set of that origin
node. The backward reachable set (or in-going contact chain) is the set
of nodes that reach a node."}`
