---
title: "IAA EPIC pigs dataflow"
output: html_notebook
---
```{r set-up, results = "hide"}
#library(movenet)
load_all()
library(parallel)
n_threads <- detectCores()-1
library(lubridate)
library(networkDynamic)
library(tidyverse)
```

## Loading in data 

### Questions

* <mark>Approximate timing on obtaining data access?</mark>
* Which additional movement data columns are useful for important network analyses? (e.g. purpose of movement?)
* Which holding data columns are useful for important network analyses? (e.g. holding size, production, purpose)
* What data checking/cleaning steps are needed for the pig movement and holding data before creating a network?
* (From how much time to )

```{r load-data}
#config_file <- "fakeScotEID_combined_config.yml"
config_file <- "ScotEID_movement.yml"
load_config(config_file)
#movement_data <- reformat_data(movement_data_file_or_connection, type = "movement") #movement data
movement_data <- example_movement_data 
#holding_data <- reformat_data(holding_data_file_or_connection, type = "holding") #holding data
#holding_data <- example_holding_data
```
### To do

* Extract holding data from ScotEID
* Create holding "ScotEID_holding.yml" or make combined config
* Create data cleaning script
  * Check duplicates
  * Check missing data for from, to, date, weight, holding id, holding production, holding purpose, herd size
  * Check holding identifier is unique in various holding data tables, if not, how do you select?

## Modifying data and creating network representations

### Questions - data modification

* <mark>What are sensible jitter ranges and rounding units to consider?</mark>
  * rounding dates: week, month, (2 or 3 months?) + Aggregate yes/no?
  * jittering dates: +- 3, 7, 15 days? + n simulations?
  * rounding weights (batch sizes): 5, 10, 50, 100? 
  * jittering weights (batch sizes): ?
* <mark>What if anything do we want to do with hexscape geojitter function? distance/coords not part of network measures but relevant in models</mark>  
* <mark>Treat all the same or require different treatments for different types of holdings?</mark>  
* <mark>Immediately create networks for all combinations of date treatment + weight treatment, or let this depend on first analyses?</mark> 

### Questions - network generation

* <mark>The following assumptions are made in creating temporal networks: networks are 
assumed to be directed, monopartite, and with no loops allowed. Are these assumptions 
reasonable for pig data? For cattle and sheep, what would need to change?</mark>
* <mark>Include non-active holdings yes/no? How does this matter?</mark>

```{r modify-data}
jitter_dates_set <- c(3, 7, 15) #corresponding to half rounding units
jitter_dates_n_sim <- 5 #How many simulations to do for each jitter value?
round_dates_set <- c("week", "month", "bimonth")

create_jittered_dates_networks <- function(movement_data, movement_configfile, 
                                           jitter_set, n_sim, n_threads,
                                           holding_data = NULL, 
                                           incl_nonactive_holdings = FALSE){

  cl <- makeCluster(n_threads)
  on.exit(stopCluster(cl))
  clusterExport(cl, c("movement_data", "holding_data", "movement_configfile", 
                      "incl_nonactive_holdings"), envir = environment())
  clusterExport(cl, c("has_element", "floor_date", "select", "networkDynamic"))
  clusterEvalQ(cl, {
    library(movenet)
    library(checkmate)
    library(dplyr)
    load_config(movement_configfile)
  })
  jitter_networks <-
    pblapply(rep(jitter_set, n_sim),
             function(x){jitter_dates(movement_data, range = x) |>
                         movedata2networkDynamic(holding_data = holding_data,
                                                 incl_nonactive_holdings = incl_nonactive_holdings)},
             cl = cl)
  names(jitter_networks) <- paste0("jitter (",rep(jitter_set, n_sim)," days)")
  return(jitter_networks)
}

create_rounded_dates_networks <- function(movement_data, movement_configfile, 
                                          round_set, n_threads, holding_data = NULL, 
                                          incl_nonactive_holdings = FALSE,
                                     week_start = getOption("lubridate.week.start", 7),
                                     sum_weight = TRUE, ...){
  cl <- makeCluster(n_threads)
  on.exit(stopCluster(cl))
  clusterExport(cl, c( "week_start", "has_element", "floor_date", "select", "networkDynamic"))
  clusterExport(cl, c("movement_data", "holding_data", "movement_configfile", 
                      "incl_nonactive_holdings"), envir = environment())
  clusterEvalQ(cl, {
    library(movenet)
    library(checkmate)
    library(dplyr)
    load_config(movement_configfile)
  })
  rounding_networks <-
    pblapply(round_set,
             function(x){round_dates(movement_data, unit = x,
                                     week_start = week_start,
                                     sum_weight = sum_weight, ...) |>
                         movedata2networkDynamic(holding_data = holding_data, 
                                                 incl_nonactive_holdings = 
                                                   incl_nonactive_holdings)},
             cl = cl)
  names(rounding_networks) <- paste0(round_set,"ly")
  return(rounding_networks)
}

jitter_networks <- 
  create_jittered_dates_networks(movement_data, config_file, jitter_dates_set, 
                                 jitter_dates_n_sim, n_threads, holding_data, 
                                 incl_nonactive_holdings = FALSE)
week_start <- wday(min(movement_data[[movenet:::movenetenv$options$movedata_cols$date]]))
rounding_networks <- 
  create_rounded_dates_networks(movement_data, config_file, round_dates_set, 
                                n_threads, holding_data, 
                                incl_nonactive_holdings = FALSE, 
                                week_start = week_start, sum_weight = TRUE)

```


### To do
*
* (Remove assumptions by setting additional arguments or using '...')

